---
title: "ANOVA"
output:
  pdf_document: default
  html_document: default
date: "2024-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
Describe how you will conduct a factorial experiment with two factors at a and b levels as an RCBD when there is one nuisance factor at c levels. Write down the model, assumptions and ANOVA table.\

To conduct a factorial experiment with two factors at a and b levels as an RCBD when there is one nuisance factor at c levels, I identify the model as $y_{ijk}=\mu+\tau_i+\beta_j+(\tau\beta)_{ij}+\epsilon_{ijk}$ where $i=1,...,a$; $j=1,...,b$; and $k=1,...,n$. There are $a$ levels of Factor A, $b$ levels for Factor B, and $n$ is replicates within each combination of levels. Because this is a randomized complete block design (RCBD), we assume the experimental conditions are homogenous and there are no other important effects beyond the experimental factor. To deal with the nuisance factor at $c$ levels, we use the technique of blocking since we know from the beginning that factor $c$ has effects we aren't interested in. The ANOVA table for factorial design is below (note all equations to determine F value have the denominator of MSE).\
```{r}
#install.packages("readxl")
library(readxl)
anovaQ1 <- read_excel(path="~/Desktop/PSTAT122/Data/HW6Q1.xlsx")
print(anovaQ1)
```

### Question 2
The following output was obtained from a computer program that performed a two-way ANOVA on a factorial experiment.\

#### a
Fill in the blanks in the ANOVA table.\
```{r}
anovaQ2 <- read_excel(path="~/Desktop/PSTAT122/Data/HW6Q2.xlsx")
print(anovaQ2)
```

#### b
How many levels were used for Factor B?\
DF of interaction:\
$(a-1)(b-1)=3$\
$(2-1)(b-1)=3$\
$b-1=3$\
$b=4$\
Four levels were used for Factor B.\

#### c
How many replicates of the experiment were performed?\
DF of total:\
$abn-1=15$\
$2*4*n-1=15$\
$8n=16$\
$n=2$\
Two replicates of the experiment were performed.\

#### d
What conclusions would you draw?\
Based on the ANOVA table, I can conclude that there is no significant difference amongst the two factor main effects or the interaction at the 95% significance level.\

### Question 3
Consider the battery life experiment (example 5.1 of text, pg 161) discussed in class. note that the interaction is significant. For material type 1, conduct a pairwise comparison of the mean battery life under the three temperature settings using Fisher LSD method. Now for temperature=125F, conduct a pairwise comparison of the mean battery life under the three material types using Fisher LSD method.\
```{r}
library(agricolae)
values <- c(130,115,74,180,
            34,40,80,75,
            20,70,82,58)
data_frame <- data.frame(life=values,temp=factor(rep(1:3,each=4)))
fit <- aov(life~temp,data=data_frame)
anova(fit)
test1 <- LSD.test(fit,"temp")
test1

values <- c(20,70,82,58,
            25,70,58,45,
            96,104,82,60)
data_frame = data.frame(life=values,mat=factor(rep(1:3,each=4)))
fit <- aov(life~mat, data=data_frame)
anova(fit)
test2 <- LSD.test(fit, "mat")
test2
```
Based on the above tests, we conclude a significant difference in average life between the 15`F temperature and both the 70`F and 125`F. The test evaluating differences within 125`F found a significant difference between material 1 and material 3.\

### Question 4
The factors that influence the breaking strength of a synthetic fiber are
being studied. Four production machines and three operators are chosen
and a factorial experiment is run using fiber from the same production
batch. The results are as follows.\

```{r}
fiber_strength <- c(109,110,108,110,
                    110,115,109,108,
                    110,110,111,114,
                    112,111,109,112,
                    116,112,114,120,
                    114,115,119,117)
operators <- c(rep("1",8), rep("2",8), rep("3",8))
machines <- factor(c(1,2,3,4,1,2,3,4,1,2,3,4,1,2,3,4,1,2,3,4,1,2,3,4))
data_frame <- data.frame(fiber_strength,machines,operators)
```

#### a
Analyze the data and draw conclusions.\
```{r}
fit <- aov(fiber_strength~machines+operators+machines:operators,data=data_frame)
summary(fit)
```
Based on the summary above, we can note that, if evaluating at the 95% significance level, there is a significant difference among at least one of the operators, as the p-value is for operators is smaller than 0.05.\

#### b
Prepare appropriate residual plots and comment on the validity of model assumptions.\
```{r}
resi_duals <- residuals(fit)
qqnorm(resi_duals)
qqline(resi_duals)
```
Based on the above plot, we can conclude that the data does not follow a normal distribution. As the plot points do not resemble a line.\